Model-View-Controller Package

### 2006-05-05 inst/doc/MVCClass.Rnw
Added some new classes and changed a few existing classes.  Plus had to change all the documentation.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MVCClass@17877 bc3139a8-67e5-0310-9ffc-ced21a209358

\title{Model-View-Controller Package}
\author{Elizabeth Whalen}
\date{April 20, 2006}

@@

The \Robject{MVC} class represents a model-view-controller object.  Basically, 
this class will bind the model with its views and its controller so that 
an object of this class can be created each time there is a new data set 
(model).  Thus, there is a one-to-one relationship between a MVC object and
a model object.  Because of this relationship, the
MVC object and the model object are identified by the same name,
which is stored in the modelName slot of the model object (see Section 
\ref{Sec:Model}).  The figure below 
shows the class definition for \Robject{MVC}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.3}{\includegraphics{MVCClass.jpg}}
    \caption{ The MVC Class. }
    \label{Fig:MVCClass}
  \end{center}
\end{figure}

The \Robject{MVC} class has five slots which are model, viewList, controller, 
parentMVC, and childMVCList.  The
model slot stores the model object, the viewList slot is a list of the view
objects for this model, the controller slot is an environment that stores
information for this MVC, the parentMVC slot is the name of the parent model
(MVC), and the childMVCList slot is a list of the children models
(MVCs).  Note that model and MVC objects use the same name so the
parentMVC slot refers to both the name of the model and the name of the
MVC.  Similarly, for the childMVCList slot, the names in the list
refer to both the name of the model and the name of the MVC.

Thus, the \Robject{MVC} class not only binds the model, view, and controller
objects together, but MVC objects can be related to each through a 
parent-child (tree) hierarchy because of the slots, parentMVC and 
childMVCList. 
These two slots also show that a MVC object can have at most one parent MVC and
can have as many children MVCs as it wants.

@@

The above figure shows the inheritance structure for message objects that can 
be passed within one MVC object and a later figure in this section will show 
the inheritance structure for messages objects that can be passed between MVC 
objects.

@@

Currently, there are two types of messages for messaging within a model, view,
controller object: an add and an update message.  Both the add message,
\Robject{gAddMessage}, and the update message, \Robject{gUpdateMessage},
contain three slots, dataName, mData, and type.  dataName is the character
string that gives the name of the model, mData is a list of data needed to
perform the addition or alteration operation, and type is a character string
that gives the type of addition or alteration to perform.  Both
\Robject{gAddMessage} and \Robject{gUpdateMessage} are virtual classes.

Since adding a view requires different information and methods from adding a
model (and similarly with updating views versus models), it makes sense to
have two separate message classes for these operations.  Thus, when messages
are sent between components, they are from one of the following classes:
\Robject{gAddDataMessage}, \Robject{gAddViewMessage},
\Robject{gUpdateDataMessage}, and \Robject{gUpdateViewMessage}.  As shown in
Figure \ref{Fig:Mess}, the \Robject{gAddDataMessage} and
\Robject{gAddViewMessage} classes inherit from the \Robject{gAddMessage}
class, and the \Robject{gUpdateDataMessage} and \Robject{gUpdateViewMessage}
classes inherit from the \Robject{gUpdateMessage} class.

@@

\clearpage

@@

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.3}{\includegraphics{EventClass.jpg}}
    \caption{ The gEventFun Class. }
    \label{Fig:EventFun}
  \end{center}
\end{figure}

As shown in Fig \ref{Fig:EventFun}, the slots in the \Robject{gEventFun} class
are callFun, shortName, and preprocessFun.  The callFun slot is the name of
the callback function, the shortName slot is a short description of the
function, and the preprocessFun slot is a character vector of all
preprocessing functions that must be called before the callback function.

@@

The preprocessing functions that are stored in the preprocessFun slot will
take no parameters.  The purpose of the preprocessing functions is to set some
variables that the callback function will need.  Thus, not all callback
functions will need preprocessing functions and if no preprocessing functions
are needed, then this slot will be NULL.  If the callback function sets the
color of a node in a graph, then a preprocessing function is needed to
determine what color this will be.  So in this example, the preprocessing
function will open a color browser to allow the user to choose what the new
color will be.

>>>>

\title{Model-View-Controller Package}
\author{Elizabeth Whalen}
\date{May 5, 2006}

@@

The \Robject{MVC} class represents a model-view-controller object.  Basically, 
this class will bind the model with its views and its controller so that 
an object of this class can be created each time there is a new data set 
(model).  Thus, there is a one-to-one relationship between a MVC object and
a model object.  Because of this relationship, the
MVC object and the model object are identified by the same name,
which is stored in the modelName slot of the model object (see Section 
\ref{Sec:Model}).  Figure \ref{Fig:MVCClass} 
shows the inheritance structure for the \Robject{MVC} classes.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.2}{\includegraphics{fullMVCClass.jpg}}
    \caption{ Inheritance for MVC Objects. }
    \label{Fig:MVCClass}
  \end{center}
\end{figure}

The \Robject{MVC} class is a virtual class that binds the other \Robject{MVC}
classes together.  Every \Robject{MVC} object will need the three slots:
model, viewList and controller.  The model slot stores the model object, the
viewList slot is a list of the view objects for this model, and the controller
slot is an environment that stores information for this MVC.  Inheriting from
the \Robject{MVC} class is the \Robject{singleModelMVC} class that has no new
slots.  The \Robject{singleModelMVC} class represents a \Robject{MVC} object
that is not linked to any other \Robject{MVC} objects.

Inheriting from the \Robject{singleModelMVC} class is the
\Robject{linkedModelMVC} that contains the new slots: parentMVC and
childMVCList.  The parentMVC slot is the name of the parent model (MVC) and
the childMVCList slot is a list of the children models (MVCs).  Note that
model and MVC objects use the same name so the parentMVC slot refers to both
the name of the model and the name of the MVC.  Similarly, for the
childMVCList slot, the names in the list refer to both the name of the model
and the name of the MVC.

Thus, the \Robject{singleModelMVC} class binds the model, view, and controller
objects together and the \Robject{linkedModelMVC} class allows the MVC objects
to be related to each through a parent-child (tree) hierarchy because of the
slots, parentMVC and childMVCList.  These two slots also show that a MVC
object can have at most one parent MVC and can have as many children MVCs as
it wants.

@@

The figure shows the inheritance structure for message objects that can 
be passed within one MVC object and a later figure in this section will show 
the inheritance structure for messages objects that can be passed between MVC 
objects.

@@

Currently, there are two types of messages for messaging within a model, view,
controller object: an add and an update message, and both of these messages
are performing some type of modification.  Thus, both the add message,
\Robject{gAddMessage}, and the update message, \Robject{gUpdateMessage},
inherit from the virtual class, \Robject{gModifyMessage}, which has 
three slots, dataName, mData, and type.  dataName is the character
string that gives the name of the model, mData is a list of data needed to
perform the addition or alteration operation, and type is a character string
that gives the type of addition or alteration to perform.  Both
\Robject{gAddMessage} and \Robject{gUpdateMessage} are virtual classes.

Since adding a view requires different information and methods from adding a
model (and similarly with updating views versus models), it makes sense to
have two separate message classes for these elements.  Thus, when messages
are sent between components, they are from one of the following classes:
\Robject{gAddDataMessage}, \Robject{gAddViewMessage},
\Robject{gUpdateDataMessage}, and \Robject{gUpdateViewMessage}.  As shown in
Figure \ref{Fig:Mess}, the \Robject{gAddDataMessage} and
\Robject{gAddViewMessage} classes inherit from the \Robject{gAddMessage}
class, and the \Robject{gUpdateDataMessage} and \Robject{gUpdateViewMessage}
classes inherit from the \Robject{gUpdateMessage} class.

@@

%\clearpage

@@

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.2}{\includegraphics{EventClass.jpg}}
    \caption{ The gEventFun Class. }
    \label{Fig:EventFun}
  \end{center}
\end{figure}

As shown in Fig \ref{Fig:EventFun}, the slots in the \Robject{gEventFun} class
are callFun, shortName, and preprocessFun.  The callFun slot is the callback
function, the shortName slot is a short description of the 
function, and the preprocessFun slot is a list of all
preprocessing functions that must be called before the callback function.

@@

The preprocessing functions that are stored in the preprocessFun slot will
take no parameters.  The purpose of the preprocessing functions is to set some
variables that the callback function will need.  Thus, not all callback
functions need preprocessing functions and if no preprocessing functions
are needed, then this slot is set to NULL.  If the callback function sets the
color of a node in a graph, then a preprocessing function is needed to
determine what color this will be.  So in this example, the preprocessing
function opens a color browser to allow the user to choose what the new
color is.
